The Birth of a Galaxy: Population III Metal Enrichment and Population II Stellar Populations

### 2010-10-19 science_ms.tex
Adding first pass at science manuscript

\documentclass[apjl]{emulateapj}
%\documentclass[12pt,preprint]{aastex}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{natbib}
\usepackage[section] {placeins}
\bibliographystyle{apj}
\citestyle{apj}

%%%%%%%% Begin custom definitions %%%%%%%%%%%%%

\input macros.tex

%%%%%%%% End custom definitions %%%%%%%%%%

@@

\shorttitle{THE BIRTH OF A GALAXY}
\shortauthors{WISE ET AL.}

\author{John H. Wise\altaffilmark{1,2}, 
  Michael L. Norman\altaffilmark{3},
  Tom Abel\altaffilmark{4},
  Matthew J. Turk\altaffilmark{3}}

\altaffiltext{1}{Department of Astrophysical Sciences, Princeton
  University, Peyton Hall, Ivy Lane, Princeton, NJ 08544}
\altaffiltext{2}{Hubble Fellow}
\altaffiltext{3}{Center for Astrophysics and Space Sciences,
  University of California at San Diego, La Jolla, CA 92093}
\altaffiltext{4}{Kavli Institute for Particle Astrophysics and
  Cosmology, Stanford University, Menlo Park, CA 94025}
\email{jwise@astro.princeton.edu}

\begin{abstract}

  Population III stars first form in dark matter halos with masses
  around $10^6 \Ms$.  By definition, they are metal-free, and their
  protostellar collapse is driven by molecular hydrogen cooling in the
  gas-phase, leading to a massive characteristic mass $\sim 100~\Ms$
  and suppressed fragmentation.  Population II stars with lower
  characteristic masses form when the star-forming gas reaches a
  critical metallicity of $10^{-6} - 10^{-3.5}~Z_\odot$, depending on
  whether dust cooling is important.  We present adaptive mesh
  refinement radiation hydrodynamics simulations that follows the
  transition from Population III to II star formation.  We model
  stellar radiative feedback with adaptive ray tracing.  A top-heavy
  initial mass function for the Population III stars is considered,
  resulting in a plausible distribution of pair-instability supernovae
  and associated metal enrichment.  We find that the gas fraction
  recovers from 5 percent to nearly the cosmic fraction in halos with
  merger histories rich in halos above $10^7 \Ms$.  A single
  pair-instability supernova is sufficient to enrich the host halo to
  a metallicity floor of $10^{-3} Z_\odot$ and to transition to
  Population II star formation.  Neighboring halos are enriched to
  metallicities 10--100 times lower as the enriching blastwaves mainly
  bypass them.  Stellar metallicities trace gas metallicities within
  0.2 dex.  We find that stellar metallicities do not necessarily
  trace stellar ages, as mergers of halos with established stellar
  populations can create superpositions of $t-Z$ evolutionary tracks.
  A bimodal metallicity distribution is created after a starburst
  occurs when the halo can cool efficiently through atomic line
  cooling.

\end{abstract}

\keywords{cosmology --- methods: numerical --- hydrodynamics ---
  radiative transfer --- star formation}

\section{Motivation}

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \citep{Abel00, ABN02, Bromm02_P3, Yoshida03, OShea07a}.  A
fraction of these stars enrich the surrounding intergalactic medium
(IGM) when they go supernova, which can happen in stars $\lsim40~\Ms$
in Type II supernovae (SNe) or in stars roughly between 140 \Ms~and
260 \Ms~in pair-instability SNe \citep[PISNe;][]{2002ApJ...567..532H}.
The host halo and the neighboring halos are then enriched with this
ejecta.  There exists a critical metallicity that is $\sim 10^{-6}
Z_\odot$ if dust cooling is efficient \citep{Omukai05,
  Schneider06_Frag, clark08} and $\sim 10^{-3.5} Z_\odot$ otherwise
\citep{Bromm01, 2009ApJ...691..441S}, where the gas can cool rapidly,
lowering its Jeans mass.  An intermediate characteristic mass of
$\sim10~\Ms$ can be occur if the gas cooling is suppressed to the
cosmic microwave background (CMB) temperature \citep{Larson98,
  Tumlinson07_IMF, 2009ApJ...691..441S}.  The resulting Population II
star cluster will thus have a lower characteristic stellar mass than
its metal-free progenitors.  In the local universe, stars in the Milky
Way halo and dwarf spheroidal (dSph) galaxies may be connected to
these first stellar clusters.  In the halo of the Milky Way, the most
metal-poor stars have a metallicity floor of [Z/H] = --4
\citep{Beers05}.  In dwarf spheroidal (dSph) galaxies, this floor is a
factor of 10 higher \citep{Tolstoy04, Helmi06}, however several metal
poor stars with $-4 < [\mathrm{Z/H}] < -3$ have been recently
discovered in dSphs \citep{Tafelmeyer10, Frebel10_Obs}.

The transition from Pop III to Pop II star formation (SF) is solely
dependent on the propagation of metals from the SNe remnants into
future sites of SF.  Their flows are complex because of the
interactions between the SN blastwave, cosmological accretion and halo
mergers, and nearby stellar feedback.  In minihalos ($M_{\rm h} \sim
10^6~\Ms$), radiation from the massive Pop III drives a 30 \kms~shock,
which is 10 times greater than the escape velocity of the halo, and
leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5 \cubecm)
medium \citep{Kitayama04, Whalen04, Abel07}.  This aids in the
expansion of the blastwave because it delays the transition to the
Sedov-Taylor and snowplow phases.  In PISNe, approximately half of the
metals stay in the IGM with a metal bubble size of a few kpc.  The
remaining metals fall back into the host halo.  The blastwave may
induce SF in nearby halos through the compression of the gas
\citep{Ferrara98}; however timescales for metal mixing into the dense
gas are many dynamical times \citep{Cen08} for shock velocities $\lsim
100~\kms$.

@@

\section{Method}
\label{sec:setup}

@@

\subsection{Simulation setup}

We use the adaptive mesh refinement (AMR) code
\enzo~v2.0\footnote[\dag]{\texttt{enzo.googlecode.com, unstable
    changeset b86d8ba026d6}} \citep{BryanNorman1997, OShea2004}, which
has been modified to use a HLLC Riemann solver \citep{Toro94_HLLC} for
additional stability in strong shocks.  To resolve minihalos with at
least 100 dark matter (DM) particles and follow the formation of the
first generation of dwarf galaxies, we use a simulation box of 1 Mpc
that has a resolution of $256^3$.  This gives us a DM mass resolution
of 1840 \Ms.  We refine the grid on baryon overdensities of $3 \times
2^{-0.2l}$, where $l$ is the AMR level, resulting in a
super-Lagrangian behavior.  We also refine on a DM overdensity of $3
\times 2^l$ and always resolve the local Jeans length by at least 4
cells, avoiding artificial fragmentation during gaseous collapses
\citep{Truelove97}.  We initialize the simulation with \textsl{grafic}
\citep{Bertschinger01} at $z = 130$ and use the cosmological
parameters from the 7-year WMAP data \citep{WMAP7}: $\Omega_M =
0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b = 0.0449$, $h = 0.71$,
$\sigma_8 = 0.81$, and $n = 0.963$ with the variables having their
usual definitions.  We stop the simulation at $z=7$.

We use a non-equilibrium chemistry solver with 9 species of hydrogen,
helium, and molecular hydrogen \citep{Abel97}.  We will follow-up this
study with one that considers radiative cooling from metal-enriched
gas, using the method of \citet{2008MNRAS.385.1443S}.

% We peform two simulations with the same initial conditions.  They
% differ only in that one considers primordial cooling, and the other
% considers additional cooling in metal-enriched gas.  For primordial
% cooling, we use a non-equilbrium chemistry solver with 9 species of
% hydrogen, helium, and molecular hydrogen.  For metal cooling, we use a
% cooling curve that is calculated from CLOUDY \citep{CLOUDY} and using
% the method of \citet{2008MNRAS.385.1443S}.  It is tabulated based on
% density, temperature, electron fraction, and metallicity.

\subsection{Star formation}

@@

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$.  They form in molecular clouds that
coexist with the dark matter halo center with limited fragmentation
occurring during their collapse; however \citet{2009Sci...325..601T}
and \citet{Stacy10_Binary} have recently shown that Pop III binaries
may form in a fraction of such halos.

% Furthermore, detailed one-dimensional calculations have shown that
% their IMF should follow a Kroupa-like IMF that has a Salpeter slope
% at the high-mass end and an exponential cut-off below $M_{\rm char}$
% \citep{refs}.

For Pop III stars, we use the same SF model as \citet{Abel07} and
\citet{Wise08_Gal} where each star particle represents a single star.
The critical overdensity in which a star forms is $5 \times 10^5$.
Instead of using a fixed stellar mass, we randomly sampled from an IMF
with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\citep{Chabrier03, Clark09}.  For reproducibility, we record the
number of times the random number generator \citep[Mersenne
  twister;][]{MTwister} has been called for use when restarting the
simulations.

@@

We treat Pop II SF with the same prescription as \citet{Wise09}, which
is a modified version of the \citet{Cen92} method but accounts for the
fact that the molecular clouds are resolved.  The critical overdensity
is the same as the Pop III SF model.  In each star-forming region,
seven percent of the cold gas ($T < 10^3$ K) is removed from the grid
and deposited into the star particle.  Each star particle lives for 20
Myr, the maximum lifetime of an OB star.  These stars generate the
majority of the ionizing radiation and SNe feedback in stellar
clusters, thus we ignore lower mass stars.

\subsection{Stellar feedback}

The mass-dependent luminosities and lifetimes of the Pop III stars are
taken from \citet{Schaerer02}.  The radiation field is evolved with
adaptive ray tracing \citep{Abel02_RT, Wise10} and is coupled
self-consistently to the hydrodynamics.  We model the \hh~dissociating
radiation with an optically-thin, inverse square profile, centered on
all stars.  These stars die as pair-instability SNe (PISNe) if they
are in the mass range 140--260 \Ms~\citep{Heger03}.  We use the
explosion energy from \citet{Heger02}, where we fit the following
function to their models, $E_{\rm PISN} = 10^{51} \times [5.0 + 1.304
(M_{\rm He} - 64)]$, where $M_{\rm He} = (13/24) \times (M_\star - 20)
\Ms$ is the helium core mass and $M_\star$ is the stellar mass.

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime, and we do not consider singly- and doubly-ionizing
helium photons.  We note that low-metallicity stars generate up to a
factor of four more ionizing photons than a solar metallicity
population \citep{Schaerer03} and might be underestimating the
radiative feedback.  Nonetheless this study provides an excellent
first insight in the transition to Pop II SF, as the metal enrichment
is the key ingredient here.  For SN feedback, these stars generate
$6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$ of energy after living
for 4 Myr, which is injected into spheres with a radius of 10 pc.  If
the resolution of the grid is less than 10/3 pc, we deposit the energy
into a $3^3$ cube surrounding the star particle.

\section{Results}
\label{sec:results}

\begin{figure*}
  \epsscale{0.8}
  \plotone{f3}
  \epsscale{1}
  \caption{\label{fig:projections} Density-weighted projections of gas
    density (top), temperature (middle), and metallicity (bottom) at
    $z=7$.  The left column shows the entire simulation volume, where
    the center and right columns focus on the intense and quiet halos,
    which are marked by arrows in the upper-left panel.  The
    metallicity projections are a composite picture of metals
    originating from Pop III (red) and Pop II (blue) stars with purple
    indicating a mixture of the two.}
\end{figure*}

@@

\subsection{Mass evolution}
\label{sec:halo}

\begin{figure*}
  \epsscale{1.15}
  \plottwo{f1}{f2}
  \caption{\label{fig:evo} (a) Evolution of the total halo
    mass (top), stellar mass (middle), and gas fraction (bottom) of
    the quiet (dashed) and intense (solid) halos.  (b)
    Mass-weighted stellar metallicities and gas metallicities enriched
    by Pop II and Pop III SNe of the intense (top) and quiet (bottom)
    halos.}
  \epsscale{1}
\end{figure*}

@@

Both halos start forming Pop II stars when $M = 10^7 \Ms$.  This is
consistent with the filtering mass $M_f$ of high-redshift halos when
it accretes mainly from a pre-heated IGM \citep{gnedin98, gnedin00,
  Wise08_Gal}.  Afterwards these halos can cool efficiently through
\hh~cooling, sustaining constant and sometimes bursty SF.  The latter
characteristics are equated with the definition of a galaxy.  The
quiet halo forms $10^5 \Ms$ of stars by $z=9$.  This initial starburst
photo-evaporates the majority of its molecular clouds, in addition to
heating and ionizing the surrounding IGM out to a radius of 10--15 kpc
at $z=9$.  These respectively reduce the in-situ and external cold gas
supply that could feed future SF.

@@

\subsection{Metallicity evolution}

The evolution of the stellar and gas metallicity of both halos are
illustrated in Figure \ref{fig:evo}b.  PISNe from Pop III stars enrich
the nearby IGM out to a radius of 10 kpc and provides a metallicity
floor of $[Z_3/H] \sim -3$, given $M_{\rm char} = 100 \Ms$.  The
metallicity from Pop II SNe initially enrich the ISM of both halos to
an average $[\bar{Z}_2/H]$ between --3.5 and --3.  The quiet halo is
enriched to $[\bar{Z}_2/H] \sim -2.5$ until the metal-poor inflows
balance the metal-rich galactic outflows, which are directed in the
polar directions of the gas disk, keeping the adjacent filaments
metal-poor.  These features and a well-mixed ISM
\citep[cf.][]{Wise08_Gal, Greif10} are apparent in the metallicity
projections in Figure \ref{fig:projections}.  The average stellar
metallicities is equal or slightly higher up to 0.5 dex than
$[\bar{Z}_2/H]$.

@@

\subsection{Population III star formation history}

@@

\subsection{Population II stellar populations}
\label{sec:pop}

\begin{figure*}
  \plottwo{f4a}{f4b}
  \caption{\label{fig:pops} The scatter plots show the SF history of
    the quiet (left) and intense (right) halos as a function of
    metallicity at $z=7$.  Each circle represents a star cluster,
    whose area is proportional to its mass.  The open circles in the
    upper right represent sizes of $10^3$, $10^4$, and $10^5$ \Ms~star
    clusters.  The dashed lines in the right panel guide the eye to
    two stellar populations that were formed in two satellite halos,
    merging at $z=7.5$.  The upper histogram collects the SF history
    into 25 equal temporal bins.  The right histogram depicts the
    stellar metallicity distribution.}
\end{figure*}

@@

\section{Discussion and Conclusions}

In this paper, we focus on the birth of two galaxies prior to
reionization with a cosmological AMR radiation hydrodynamics
simulation.  Supernovae from Pop III stars provide the necessary heavy
elements for the transition to a Population II stellar population,
which we have directly simulated.  These two galaxies have a 10--15\%
probability in surviving as present-day ``fossil'' galaxies
\citep{Gnedin06}, otherwise they will be incorporated into galactic
stellar halos.

>>>>

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

\usepackage{graphicx}

\newcommand\aap{{A\&A}}% 
         % Astronomy and Astrophysics 
\newcommand\aaps{{A\&AS}}% 
         % Astronomy and Astrophysics, Supplement 
\newcommand\aj{{AJ}}% 
         % Astronomical Journal 
\newcommand\araa{{ARA\&A}}% 
         % Annual Review of Astron and Astrophys 
\newcommand\apj{{ApJ}}% 
         % Astrophysical Journal 
\newcommand\apjl{{ApJ}}% 
         % Astrophysical Journal, Letters 
\newcommand\apjs{{ApJS}}% 
         % Astrophysical Journal, Supplement 
\newcommand\mnras{{MNRAS}}% 
         % Monthly Notices of the RAS 
\newcommand\nat{{Nature}}% 
         % Nature 
\newcommand\pasp{{PASP}}% 
         % Publications of the ASP 
\newcommand{\cubecm}{\ifmmode{{\rm cm^{-3}}}\else{cm$^{-3}$}\fi}
\newcommand{\lsim}{\lower0.3em\hbox{$\,\buildrel <\over\sim\,$}}
\newcommand{\gsim}{\lower0.3em\hbox{$\,\buildrel >\over\sim\,$}}
\newcommand{\kms}{\ifmmode{{\rm km~s^{-1}}}\else{km s$^{-1}$}\fi}
\newcommand{\hh}{H$_2$}
\newcommand{\Ms}{\ifmmode{M_\odot}\else{$M_\odot$}\fi}
\newcommand\tento[1]{$10^{#1}$}
\newcommand{\tvir}{\ifmmode{T_{\rm{vir}}}\else{$T_{\rm{vir}}$}\fi}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

%% We will uncomment this later
%\usepackage{scicite}
\bibliographystyle{Science}

% Use times if you have the font installed; otherwise, comment out the
% following line.

\usepackage{times}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{John H.~Wise,$^{1,2\ast}$ Michael L.~Norman,$^{3}$ Tom Abel$^{4}$,
Matthew J.~Turk$^{3}$\\
\\
\normalsize{$^{1}$Department of Astrophysical Sciences, Princeton
University,}\\
\normalsize{Peyton Hall, Ivy Lane, Princeton, NJ 08544}\\
\normalsize{$^{2}$Hubble Fellow}\\
\normalsize{$^{3}$Center for Astrophysics and Space Sciences,}\\
\normalsize{University of California at San Diego, La Jolla, CA 92093}\\
\normalsize{$^{4}$Kavli Institute for Particle Astrophysics and
  Cosmology, }\\
\normalsize{Stanford University, Menlo Park, CA 94025}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:
jwise@astro.princeton.edu}
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

@@

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
  Population III stars first form in dark matter halos with masses
  around $10^6~\Ms$.  By definition, they are metal-free, and their
  protostellar collapse is driven by molecular hydrogen cooling in the
  gas-phase, leading to a massive characteristic mass $\sim 100~\Ms$
  and suppressed fragmentation.  Population II stars with lower
  characteristic masses form when the star-forming gas reaches a
  critical metallicity of $10^{-6} - 10^{-3.5}~Z_\odot$, depending on
  whether dust cooling is important.  We present adaptive mesh
  refinement radiation hydrodynamics simulations that follows the
  transition from Population III to II star formation.  We model
  stellar radiative feedback with adaptive ray tracing.  A top-heavy
  initial mass function for the Population III stars is considered,
  resulting in a plausible distribution of pair-instability supernovae
  and associated metal enrichment.  We find that the gas fraction
  recovers from 5 percent to nearly the cosmic fraction in halos with
  merger histories rich in halos above $10^7~\Ms$.  A single
  pair-instability supernova is sufficient to enrich the host halo to
  a metallicity floor of $10^{-3}~Z_\odot$ and to transition to
  Population II star formation.  Neighboring halos are enriched to
  metallicities 10--100 times lower as the enriching blastwaves mainly
  bypass them.  Stellar metallicities trace gas metallicities within
  0.2 dex.  We find that stellar metallicities do not necessarily
  trace stellar ages, as mergers of halos with established stellar
  populations can create superpositions of $t-Z$ evolutionary tracks.
  A bimodal metallicity distribution is created after a starburst
  occurs when the halo can cool efficiently through atomic line
  cooling.
\end{sciabstract}

% In setting up this template for *Science* papers, we've used both
% the \section* command and the \paragraph* command for topical
% divisions.  Which you use will of course depend on the type of paper
% you're writing.  Review Articles tend to have displayed headings, for
% which \section* is more appropriate; Research Articles, when they have
% formal topical divisions at all, tend to signal them with bold text
% that runs into the paragraph, for which \paragraph* is the right
% choice.  Either way, use the asterisk (*) modifier, as shown, to
% suppress numbering.

\section*{Motivation}

The first (Pop III) stars are metal-free and have a large
characteristic mass and suppressed fragmentation in its protostellar
collapse \cite{Abel00, ABN02, Bromm02_P3, Yoshida03, OShea07a}.  A
fraction of these stars enrich the surrounding intergalactic medium
(IGM) when they go supernova, which can happen in stars $\lsim 40~\Ms$
in Type II supernovae (SNe) or in stars roughly between 140 \Ms~and
260 \Ms~in pair-instability SNe (PISNe; \cite{2002ApJ...567..532H}).
The host halo and the neighboring halos are then enriched with this
ejecta.  There exists a critical metallicity that is $\sim 10^{-6}
Z_\odot$ if dust cooling is efficient \cite{Omukai05,
  Schneider06_Frag, clark08} and $\sim 10^{-3.5} Z_\odot$ otherwise
\cite{Bromm01, 2009ApJ...691..441S}, where the gas can cool rapidly,
lowering its Jeans mass.  An intermediate characteristic mass of
$\sim10~\Ms$ can be occur if the gas cooling is suppressed to the
cosmic microwave background (CMB) temperature \cite{Larson98,
  Tumlinson07_IMF, 2009ApJ...691..441S}.  The resulting Population II
star cluster will thus have a lower characteristic stellar mass than
its metal-free progenitors.  In the local universe, stars in the Milky
Way halo and dwarf spheroidal (dSph) galaxies may be connected to
these first stellar clusters.  In the halo of the Milky Way, the most
metal-poor stars have a metallicity floor of [Z/H] = --4
\cite{Beers05}.  In dwarf spheroidal (dSph) galaxies, this floor is a
factor of 10 higher \cite{Tolstoy04, Helmi06}, however several metal
poor stars with $-4 < [\mathrm{Z/H}] < -3$ have been recently
discovered in dSphs \cite{Tafelmeyer10, Frebel10_Obs}.

The transition from Pop III to Pop II star formation (SF) is solely
dependent on the propagation of metals from the SNe remnants into
future sites of SF.  Their flows are complex because of the
interactions between the SN blastwave, cosmological accretion and halo
mergers, and nearby stellar feedback.  In minihalos ($M_{\rm h} \sim
10^6~\Ms$), radiation from the massive Pop III drives a 30 \kms~shock,
which is 10 times greater than the escape velocity of the halo, and
leaves behind a warm ($3 \times 10^4$ K) and diffuse (0.5 \cubecm)
medium \cite{Kitayama04, Whalen04, Abel07}.  This aids in the
expansion of the blastwave because it delays the transition to the
Sedov-Taylor and snowplow phases.  In PISNe, approximately half of the
metals stay in the IGM with a metal bubble size of a few kpc.  The
remaining metals fall back into the host halo.  The blastwave may
induce SF in nearby halos through the compression of the gas
\cite{Ferrara98}; however timescales for metal mixing into the dense gas are
many dynamical times \cite{Cen08} for shock velocities $\lsim100~\kms$.

@@

\section*{Method}
\label{sec:setup}

@@

\subsection*{Simulation setup}

We use the adaptive mesh refinement (AMR) code
enzo~v2.0\footnote{\texttt{enzo.googlecode.com, unstable changeset
b86d8ba026d6}} \cite{BryanNorman1997, OShea2004}, which has been modified to
use a HLLC Riemann solver \cite{Toro94_HLLC} for additional stability in strong
shocks.  To resolve minihalos with at least 100 dark matter (DM) particles and
follow the formation of the first generation of dwarf galaxies, we use a
simulation box of 1 Mpc that has a resolution of $256^3$.  This gives us a DM
mass resolution of 1840 \Ms.  We refine the grid on baryon overdensities of $3
\times 2^{-0.2l}$, where $l$ is the AMR level, resulting in a super-Lagrangian
behavior.  We also refine on a DM overdensity of $3 \times 2^l$ and always
resolve the local Jeans length by at least 4 cells, avoiding artificial
fragmentation during gaseous collapses \cite{Truelove97}.  We initialize the
simulation with \textsl{grafic} \cite{Bertschinger01} at $z = 130$ and use the
cosmological parameters from the 7-year WMAP data \cite{WMAP7}: $\Omega_M =
0.266$, $\Omega_\Lambda = 0.734$, $\Omega_b = 0.0449$, $h = 0.71$, $\sigma_8 =
0.81$, and $n = 0.963$ with the variables having their usual definitions.  We
stop the simulation at $z=7$.

We use a non-equilibrium chemistry solver with 9 species of hydrogen,
helium, and molecular hydrogen \cite{Abel97}.  We will follow-up this
study with one that considers radiative cooling from metal-enriched
gas, using the method of \cite{2008MNRAS.385.1443S}.

% We peform two simulations with the same initial conditions.  They
% differ only in that one considers primordial cooling, and the other
% considers additional cooling in metal-enriched gas.  For primordial
% cooling, we use a non-equilbrium chemistry solver with 9 species of
% hydrogen, helium, and molecular hydrogen.  For metal cooling, we use a
% cooling curve that is calculated from CLOUDY \cite{CLOUDY} and using
% the method of \cite{2008MNRAS.385.1443S}.  It is tabulated based on
% density, temperature, electron fraction, and metallicity.

\subsection*{Star formation}

@@

Simulations have shown that the characteristic mass of Pop III stars
$M_{\rm char} \sim 100~\Ms$.  They form in molecular clouds that
coexist with the dark matter halo center with limited fragmentation
occurring during their collapse; however \cite{2009Sci...325..601T}
and \cite{Stacy10_Binary} have recently shown that Pop III binaries
may form in a fraction of such halos.

% Furthermore, detailed one-dimensional calculations have shown that
% their IMF should follow a Kroupa-like IMF that has a Salpeter slope
% at the high-mass end and an exponential cut-off below $M_{\rm char}$
% \cite{refs}.

For Pop III stars, we use the same SF model as \cite{Abel07} and
\cite{Wise08_Gal} where each star particle represents a single star.
The critical overdensity in which a star forms is $5 \times 10^5$.
Instead of using a fixed stellar mass, we randomly sampled from an IMF
with a functional form of
%
\begin{equation}
\label{eqn:imf}
f(M)dM = M^{-1.3} \exp\left[-\left(\frac{M_{\rm char}}{M}\right)^{1.6}\right]
\end{equation}
to determine the stellar mass.  Above $M_{\rm char}$, it behaves as a
Salpeter IMF but is exponentially cutoff below that mass
\cite{Chabrier03, Clark09}.  For reproducibility, we record the
number of times the random number generator (Mersenne twister; \cite{MTwister})
has been called for use when restarting the simulations.

@@

We treat Pop II SF with the same prescription as \cite{Wise09}, which
is a modified version of the \cite{Cen92} method but accounts for the
fact that the molecular clouds are resolved.  The critical overdensity
is the same as the Pop III SF model.  In each star-forming region,
seven percent of the cold gas ($T < 10^3$ K) is removed from the grid
and deposited into the star particle.  Each star particle lives for 20
Myr, the maximum lifetime of an OB star.  These stars generate the
majority of the ionizing radiation and SNe feedback in stellar
clusters, thus we ignore lower mass stars.

\subsection*{Stellar feedback}

The mass-dependent luminosities and lifetimes of the Pop III stars are
taken from \cite{Schaerer02}.  The radiation field is evolved with
adaptive ray tracing \cite{Abel02_RT, Wise10} and is coupled
self-consistently to the hydrodynamics.  We model the \hh~dissociating
radiation with an optically-thin, inverse square profile, centered on
all stars.  These stars die as pair-instability SNe (PISNe) if they
are in the mass range 140--260 \Ms~\cite{Heger03}.  We use the
explosion energy from \cite{Heger02}, where we fit the following
function to their models, $E_{\rm PISN} = 10^{51} \times [5.0 + 1.304
(M_{\rm He} - 64)]$, where $M_{\rm He} = (13/24) \times (M_\star - 20)
\Ms$ is the helium core mass and $M_\star$ is the stellar mass.

The Pop II stars emit 6000 hydrogen ionizing photons per baryon over
their lifetime, and we do not consider singly- and doubly-ionizing
helium photons.  We note that low-metallicity stars generate up to a
factor of four more ionizing photons than a solar metallicity
population \cite{Schaerer03} and might be underestimating the
radiative feedback.  Nonetheless this study provides an excellent
first insight in the transition to Pop II SF, as the metal enrichment
is the key ingredient here.  For SN feedback, these stars generate
$6.8 \times 10^{48}$ erg s$^{-1}$ $\Ms^{-1}$ of energy after living
for 4 Myr, which is injected into spheres with a radius of 10 pc.  If
the resolution of the grid is less than 10/3 pc, we deposit the energy
into a $3^3$ cube surrounding the star particle.

\section*{Results}
\label{sec:results}

@@

\subsection*{Mass evolution}
\label{sec:halo}

@@

Both halos start forming Pop II stars when $M = 10^7 \Ms$.  This is
consistent with the filtering mass $M_f$ of high-redshift halos when
it accretes mainly from a pre-heated IGM \cite{gnedin98, gnedin00,
  Wise08_Gal}.  Afterwards these halos can cool efficiently through
\hh~cooling, sustaining constant and sometimes bursty SF.  The latter
characteristics are equated with the definition of a galaxy.  The
quiet halo forms $10^5 \Ms$ of stars by $z=9$.  This initial starburst
photo-evaporates the majority of its molecular clouds, in addition to
heating and ionizing the surrounding IGM out to a radius of 10--15 kpc
at $z=9$.  These respectively reduce the in-situ and external cold gas
supply that could feed future SF.

@@

\subsection*{Metallicity evolution}

The evolution of the stellar and gas metallicity of both halos are
illustrated in Figure \ref{fig:evo}b.  PISNe from Pop III stars enrich
the nearby IGM out to a radius of 10 kpc and provides a metallicity
floor of $[Z_3/H] \sim -3$, given $M_{\rm char} = 100 \Ms$.  The
metallicity from Pop II SNe initially enrich the ISM of both halos to
an average $[\bar{Z}_2/H]$ between --3.5 and --3.  The quiet halo is
enriched to $[\bar{Z}_2/H] \sim -2.5$ until the metal-poor inflows
balance the metal-rich galactic outflows, which are directed in the
polar directions of the gas disk, keeping the adjacent filaments
metal-poor.  These features and a well-mixed ISM
(cf.~\cite{Wise08_Gal, Greif10}) are apparent in the metallicity
projections in Figure \ref{fig:projections}.  The average stellar
metallicities is equal or slightly higher up to 0.5 dex than
$[\bar{Z}_2/H]$.

@@

\subsection*{Population III star formation history}

@@

\subsection*{Population II stellar populations}
\label{sec:pop}

@@

\section*{Discussion and Conclusions}

In this paper, we focus on the birth of two galaxies prior to
reionization with a cosmological AMR radiation hydrodynamics
simulation.  Supernovae from Pop III stars provide the necessary heavy
elements for the transition to a Population II stellar population,
which we have directly simulated.  These two galaxies have a 10--15\%
probability in surviving as present-day ``fossil'' galaxies
\cite{Gnedin06}, otherwise they will be incorporated into galactic
stellar halos.

We find that one PISN is sufficient to enrich the star-forming halo
and surrounding $\sim 5$ kpc to a metallicity of 10$^{-3} Z_\odot$,
given a Pop III characteristic mass of 100~\Ms.  If the first stars
have a lower characteristic mass that favor hypernovae
\cite{Tumlinson07_IMF}, then this metallicity floor should be lowered
by a factor of $\sim 10$ because (1) the metal ejecta is lowered by a
factor of $\sim 50$ and (2) the mixing mass is approximately decreased
by a factor of $(E_{\rm hyp}/E_{\rm PISN})^{3/5}$ in the Sedov-Taylor
solution, where $(E_{\rm hyp}/E_{\rm PISN}) \sim 0.1$ is the ratio of
explosion energies of a hypernova and PISN.  In the case where this
metallicity floor is less than the critical metallicity, then the next
instance of SF will further enrich the ISM, solidifying the transition
to Pop II SF.  We conclude that it only takes one, at most two, SNe
from Pop III stars in the halo progenitors to complete the transition
to Population II \cite{Frebel10}.  The question of
whether the critical metallicity is \tento{-6} or \tento{-3.5}
$Z_\odot$ is most applicable to nearby halos where the heavy elements
mix slowly into dense cores as the blastwave overtakes it.  Less than
two percent of stars have [Z/H] $< -3$ in both systems, consistent
with observations of metal-poor stars in the halo and dSphs
\cite{Beers05, Battaglia10}.

We also find that the merger history plays an important role in
supplying gas into halos after its first epoch of SF.  Mergers of
halos below the filtering mass are inefficient in providing gas
whereas the opposite is true for merging halos above the filtering
mass.  Halos do not necessarily need $\tvir \ge 10^4$ K to form
significant stellar populations; however SFRs dramatically increase,
and thus metal enrichment, when this threshold is reached.  Halos with
mostly gas-poor mergers or a quiet merger history result in a
monotonic increase in metallicity with time.  SFHs become more complex
with multiple metallicity-age tracks if the halo experiences mergers
with other halos that have an established stellar population.
Furthermore SF that is triggered by blastwaves interacting with
molecular clouds can have metallicities up to a factor of 10 higher
than the main starburst.  Lastly we find that a starburst at $\tvir =
10^4$ K enriches the host halo enough to create a bimodal metallicity
distribution, where the metal-poor component is created before the
burst.  Note that mergers of stellar populations can also create a
similar bimodal distribution.  Such distributions are observed in
dSphs \cite{Battaglia10}.
